---
title: "Customer Spending Analysis"
---

```{r setup, include=FALSE}
library(ggplot2)
library(knitr)
library(scales)

# Set dark theme
theme_set(theme_minimal() +
  theme(
    plot.background = element_rect(fill = "#222222"),
    panel.background = element_rect(fill = "#333333"),
    text = element_text(color = "white"),
    axis.text = element_text(color = "white")
  ))

ecomdata <- read.csv("./data/ecomdata")
```

## Data Overview

```{r data-overview}
str(ecomdata)
summary(ecomdata)
```

## Spending Patterns

```{r website-spending, fig.cap="Time on website vs spending"}
ggplot(ecomdata, aes(x=Time.on.Website, y=Yearly.Amount.Spent)) + 
  geom_point(color="orange") + 
  labs(title = "Website Engagement vs Spending",
       x = "Time on Website (minutes)",
       y = "Yearly Amount Spent ($)") +
  scale_y_continuous(labels = dollar)
```

```{r session-spending, fig.cap="Session length vs spending"}
ggplot(ecomdata, aes(x=Avg..Session.Length, y=Yearly.Amount.Spent)) + 
  geom_point(color="orange") +
  labs(title = "Session Length vs Spending",
       x = "Average Session Length (minutes)",
       y = "Yearly Amount Spent ($)") +
  scale_y_continuous(labels = dollar)
```

## Variable Relationships

```{r pairs-plot, fig.cap="Pairwise relationships"}
pairs(ecomdata[c("Avg..Session.Length", "Time.on.App", "Time.on.Website",
                "Length.of.Membership", "Yearly.Amount.Spent")],
      col = "orange",
      pch = 16,
      labels = c("Session Length", "App Time", "Website Time",
                "Membership Years", "Yearly Spend"))
```

## Membership Analysis

```{r membership-hist, fig.cap="Membership duration"}
ggplot(ecomdata, aes(x=Length.of.Membership)) + 
  geom_histogram(fill="orange", color="white", binwidth=0.5) +
  labs(title = "Customer Tenure Distribution",
       x = "Years of Membership",
       y = "Count")
```

## Predictive Modeling

```{r linear-model}
lm.fit <- lm(Yearly.Amount.Spent ~ Length.of.Membership, data=ecomdata)
summary(lm.fit)
```

```{r model-plot, fig.cap="Regression line"}
ggplot(ecomdata, aes(x=Length.of.Membership, y=Yearly.Amount.Spent)) + 
  geom_point(color="orange", alpha=0.5) +
  geom_smooth(method="lm", color="red") +
  labs(title = "Membership Duration Predicts Spending",
       x = "Years of Membership",
       y = "Annual Spending ($)") +
  scale_y_continuous(labels = dollar)
```

## Model Validation

```{r validation}
set.seed(1)
train_idx <- sample(1:nrow(ecomdata), 0.8*nrow(ecomdata))
train <- ecomdata[train_idx,]
test <- ecomdata[-train_idx,]

model <- lm(Yearly.Amount.Spent ~ Length.of.Membership, data=train)
preds <- predict(model, test)
errors <- preds - test$Yearly.Amount.Spent

metrics <- data.frame(
  RMSE = sqrt(mean(errors^2)),
  MAPE = mean(abs(errors/test$Yearly.Amount.Spent)),
  R2 = summary(model)$r.squared
)

kable(metrics, caption = "Model Performance")
```

## Conclusions

Based on the analysis of the e-commerce customer data, we can draw the following key insights:

1. **Membership Duration**: The strongest predictor of customer behavior is the length of their membership. Specifically, for every year of membership, the model indicates a coefficient of `r round(coef(lm.fit)[2], 1)` years, suggesting that longer membership durations correlate with higher engagement or purchases.

2. **Model Performance**: The linear regression model explains `r round(summary(lm.fit)$r.squared*100, 1)`% of the variance in customer behavior. This indicates that the model provides a fairly strong fit, capturing a significant portion of the factors influencing customer actions.

These findings provide valuable insights that can guide e-commerce platforms in identifying key customer segments and tailoring strategies to enhance customer retention and engagement.
